Age modifies the effect of total cholesterol level on systolic blood pressure: Analysis of evidence from the 2015 - 2018 NHANES data

### 2023-12-15 manuscript/paper.Rmd
updte
final update

---
title: "Age modifies the effect of total cholesterol level on systolic blood pressure: Analysis of evidence from the 2015 - 2018 NHANES data"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
#subtitle: "Short title goes here!"
author:
- Amos Okutse, Brown University
- Monic Colon-Vargas, Brown University
abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.

**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio. Nam consequat non ante eget aliquam. Nulla facilisis tincidunt elit. Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut. Interdum et malesuada fames ac ante ipsum primis in faucibus. Suspendisse potenti. Duis vitae nibh mauris. Duis nec sem sit amet ante dictum mattis. Suspendisse diam velit, maximus eget commodo at, faucibus et nisi. Ut a pellentesque eros, sit amet suscipit eros. Nunc tincidunt quis risus suscipit vestibulum. Quisque eu fringilla massa."
keywords: "pandoc, r markdown, knitr"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
#mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
bibliography: bibfile.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")

#source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

@@

In western societies aged 40 years and above, a demonstrable pattern
emerges, revealing an approximate elevation of 7 mmHg in SBP per decade
among individuals above 40 years. This discernible and consistent
increment in SBP with advancing age accentuates its profound impact
within the broader context of aging-related health dynamics. Similarly,
while diastolic blood pressure (DBP), demonstrates a concurrent rise
correlating with age, presents a substantially lower rate of increase
when juxtaposed against the ascending trajectory observed in SBP
[@gurven2012does; @su2022association]. Thus, in light of the
significance attached to SBP dynamics within aging populations and the
intricate interplay between age, sleep, cholestrol levels and blood
pressure alterations, our investigative focus is oriented towards the
exploration of variables that exert potential influence on systolic
blood pressure (SBP). This deliberate focus aims to elucidate the
multifaceted nature of factors contributing to SBP variability, enabling
a more comprehensive understanding of its determinants within the
framework of health and aging.

@@

Furthermore, we conducted two distinct variable selection methodologies,
similarly excluding hypertension. Model 6 was based on best subset
selection with 10-fold cross-validation to ascertain the optimal number
of subsets. Best subset selection was selected by identifying the best
model that contains a given number of predictors yielding lowest RSS.
Notably, for systolic blood pressure, the forward best subset selection
method discerned 13 variables, whereas diastolic blood pressure
exhibited 19 selected variables. However, a limitation inherent in
forward best subset selection is its inability to eliminate previously
selected features, potentially disregarding their relevance in light of
newly added variables. Consequently, this methodology may yield
sub-optimal variable selections due to its lack of adaptability.

@@

```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase. SBP = Systolic blood pressure, DBP = Diastolic blood pressure.", 
                 longtable = TRUE)
```

@@

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
Likelihood ratio tests showed a statistically significant difference
between Model 1 and 2 with evidence of in favor of the latter
($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
potentially confounding covariates in \@ref(tab:tabthree) including an
interaction between cholestrol level and age was then compared to Model
3 which adjusted for only covariates we found significant in univariate
analyses. Results here similarly showed evidence in favor of Model 2
($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a
non-statistically difference in model fit suggesting that Model 3, even
though smaller performed comparably well. Model 3 was thus selected and
compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring
its choice as the best model ($p>0.05$). Additional regression
diagnostics for Model 3 including residual versus fitted value plots,
Q-Q plots for checking the normality of residuals, Cook's distance for
outliers, residual and high leverage point plots did not reveal
substantial deviations from linear regression assumptions. Model 3 was
thus selected as the best model. Reported results are thus based on this
model.

@@

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5. These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

## Discussion and conclusion

<!-- The current study investigated the intricate interplay between age, sleep, cholestrol levels and systolic blood pressure adjusting for other potentially confounding covariates informed by literature. We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure. We find based on the best model, that work day sleep durations ranging between 6 and 8 hours and more than 8 hours are associated with a decrease in systolic blood pressure relative to the 6 hour reference group. Additionally, our study reports substantial effect modification of the effect of total cholestrol levels on systolic blood pressure by age whe adjusted for selected confounders within the study population. Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association]. -->

The present investigation explote into the settle interplay among age,
sleep patterns, cholesterol levels, and systolic blood pressure, while
accounting for various potentially confounding factors gleaned from
pertinent literature. Our study conducted an exhaustive series of model
comparisons to elucidate the intricate relationship between age, total
cholesterol levels, sleep durations, and systolic blood pressure.

In light of our findings derived from the best model based on the
performance measures in Table \@ref(tab:tabmodel), we observed that
weekday sleep durations spanning between 6 and 8 hours, as well as
durations exceeding 8 hours, exhibit a decrease in systolic blood
pressure in comparison to the reference group of individuals sleeping
for 6 hours. Furthermore, our investigation revealed a noteworthy effect
modification regarding the influence of total cholesterol levels on
systolic blood pressure when adjusted for specific confounding variables
within the studied cohort, especially in relation to age.

Our study contributes to the existing body of knowledge by
substantiating the associations uncovered in previous research
concerning sleep patterns and systolic blood pressure
[@su2022association]. This correlation underscores the significance of
sleep parameters as influential factors in understanding blood pressure
dynamics.

### Study limitations

Our study encountered a significant challenge with missing data,
resulting in a notable reduction in our sample size. Despite initially
having a sample of \$n=19,225\$, the final analysis was conducted on a
considerably smaller sample of \$n=1,977\$ observations. This
substantial reduction in the sample size raises concerns about the
robustness and reliability of our findings. Addressing this issue
becomes crucial to ensure the validity of our study outcomes.

In response to this challenge, we propose the implementation of multiple
imputation techniques. This approach involves generating multiple
plausible sets of missing values based on observed data patterns. By
imputing missing values using this method, we aim to enhance the
completeness of our data set. This strategy is intended to mitigate
potential biases resulting from missing data and fortify the reliability
and accuracy of our analytical outcomes.

>>>>

---
title: "Age modifies the effect of total cholesterol level on systolic blood pressure: Analysis of evidence from the 2015 - 2018 NHANES data"
thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
#subtitle: "Short title goes here!"
author:
- Amos Okutse$^1$
- Monica Colon-Vargas$^1$
- $^1$School of Public Health, Brown University, Providence, RI.
abstract: | 
  **Introduction:** Systolic blood pressure (SBP) is a precursor linked with a spectrum of critical health conditions.  This nexus between heightened SBP and these significant health adversities underscores the gravity of SBP as a predictive marker for adverse health outcomes.

  **Objective:** To explore the intricate interplay between SBP dynamics within non-obese aging populations with self-reported workday sleep duration, total cholesterol levels, and the total cholesterol-age dependence adjusting for selected potentially confounding factors.

  **Methods:** Cross-sectional data spanning two survey cycles (2015 – 2018) was obtained from the National Health and Nutrition Examination Survey (NHANES) database. We performed multivariable regressions to examine the interplay between age, total cholesterol levels, self-reported workday sleep duration, and SBP. Potentially confounding covariates adjusted for in analyses were selected using univariate regressions, best-subsets selection, and the Least Absolute Shrinkage and Selection Operator (LASSO). Model performance evaluation was based on information criteria, mean squared error (MSE), adjusted R-squared, and likelihood ratio tests.

  **Results:** Our results are based on a complete-case analysis of data on n = 1977 participants. We found a statistically significant association between workday sleep duration and SBP comparing $6 – 8$ hours sleep category to < $6$ hour sleep category [$\beta$ = -3.0; 95\% Confidence Interval (CI): -5.3,-0.67), p = 0.012] and a substantial effect modification of age on the effect of total cholesterol levels on SBP [$\beta$ = 0.11; 95\%CI: 0.07 - 0.14;  p <0.001].

  **Conclusion:** Workday sleep duration between 6-8 hours is associated with an optimal decrease in SBP. Age significantly modifies the effect of total cholesterol levels on SBP, adjusting for other potentially confounding factors.

  ***Keywords:*** *Systolic blood pressure*, *Sleep duration*, *Age*, *Total cholestrol level*, *Effect modification.*
keywords: "pandoc, r markdown, knitr"
#date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
bibliography: bibfile.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

```{r source-file, echo=FALSE, include=FALSE}
# source file 
#source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")

source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
```

@@

In western societies aged 40 years and above, a demonstrable pattern
emerges, revealing an approximate elevation of about 7 mmHg in SBP per decade
among individuals above 40 years. This discernible and consistent
increase in SBP with advancing age accentuates its profound impact
within the broader context of age-related health dynamics. Similarly,
while diastolic blood pressure (DBP), demonstrates a concurrent rise
correlating with age, presents a substantially lower rate of increase
when juxtaposed against the ascending trajectory observed in SBP
[@gurven2012does; @su2022association]. Thus, in light of the
significance attached to SBP dynamics within aging populations and the
intricate interplay between age, sleep, cholestrol levels and blood
pressure alterations, our investigative focus is oriented towards the
exploration of variables that exert potential influence on systolic
blood pressure (SBP). This deliberate focus aims to elucidate the
multifaceted nature of factors contributing to SBP variability, enabling
a more comprehensive understanding of its determinants within the
framework of health and aging.

@@

Furthermore, we conducted two distinct variable selection methodologies,
similarly excluding hypertension. Model 6 was based on best subset
selection with 10-fold cross-validation to ascertain the optimal number
of subsets. Best subset selection was selected by identifying the best
model that contains a given number of predictors yielding lowest RSS.
Notably, for systolic blood pressure, the forward best subset selection
method discerned 13 variables, whereas diastolic blood pressure
exhibited 19 selected variables. However, a limitation inherent in
forward best subset selection is its inability to eliminate previously
selected features, potentially disregarding their relevance in light of
newly added variables. Foward best subset deos not outperforms shringake
methods when minimizeing prediction errors [@hastie2009elements] .
Consequently, this methodology may yield sub-optimal variable selections
due to its lack of adaptability.

@@

```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
tabsum %>%
  as_kable_extra(booktabs = TRUE, 
                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender. BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase. SBP = Systolic blood pressure, DBP = Diastolic blood pressure.", 
                 longtable = TRUE) %>% 
  kable_styling(font_size = 8)
```

@@

\caption{\label{tab:tabmodel}Summary model performance measures.}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
\midrule
Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
\addlinespace
Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
\bottomrule
\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
\end{tabular}
\end{table}
```
Likelihood ratio tests showed a statistically significant difference
between Model 1 and 2 with evidence of in favor of the latter
($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
potentially confounding covariates including an
interaction between cholestrol level and age was then compared to Model
3 which adjusted for only covariates we found significant in univariate
analyses. Results here similarly showed evidence in favor of Model 2
($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a
non-statistically difference in model fit suggesting that Model 3, even
though smaller performed comparably well. Model 3 was thus selected and
compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring
its choice as the best model ($p>0.05$). Additional regression
diagnostics for Model 3 including residual versus fitted value plots,
Q-Q plots for checking the normality of residuals, Cook's distance for
outliers, residual and high leverage point plots did not reveal
substantial deviations from linear regression assumptions. Model 3 was
thus selected as the best model. Reported results are thus based on this
model.

@@

\begin{table}[H]
\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5. These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
\midrule
\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
\textbf{Sleep} &  &  &  \\
$<6$hrs &  &  &  \\
$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
\textbf{Age} & -0.04& (-0.22, 0.13) & 0.6 \\
\textbf{TC} & -3.5  & (-5.2, -1.8) & <0.001 \\
\textbf{TC*Age} & 0.11  & (0.07, 0.14) & <0.001 \\
\bottomrule
\end{tabular}
\end{table}
```
```{r tabfour, echo=FALSE, eval=FALSE}
## Final selected multiple linear regression model
tbl_regression(adj.sbp.int2,
               label = list(bmi ~ "BMI",
                            #hdl ~ "HDL",
                            total_chol ~ "TC",
                            #hemoglobin ~ "Hemoglobin",
                            albumin ~ "Albumin",
                            creatinine ~ "Creatinine",
                            #hypertension ~ "Hypertension",
                            #diabetes ~ "Diabetes",
                            citizenship_status ~ "Citizenship",
                            educ_level ~ "Education",
                            children..5 ~ "Children > 5 yrs",
                            age_yr ~ "Age (yrs)",
                            marital_status ~ "Marital status",
                            #cycle ~ "Survey cycle",
                            sleep ~ "Sleep",
                            race ~ "Race",
                            smoke ~ "Smoking status",
                            snort ~ "Snort",
                            alcohol ~ "Alcohol",
                            income_category ~ "Income",
                            gender ~ "Gender"
               )) %>% modify_header(label ~ "**Variable**") %>%
  bold_labels() %>% 
  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP). BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)

@@

# Discussion and conclusion

The present investigation explores the subtle interplay between age,
workday sleep patterns, cholesterol levels, and systolic blood pressure, while
accounting for various potentially confounding factors informed by literature. We conducted an exhaustive series of model
comparisons to elucidate the intricate relationship between age, total
cholesterol levels, workday sleep duration, and systolic blood pressure. In light of our findings derived from the best model based on the
performance measures in Table \@ref(tab:tabmodel), we observed that
weekday sleep durations spanning between 6 and 8 hours, as well as
durations exceeding 8 hours, exhibit a decrease in systolic blood
pressure in comparison to the reference group of individuals sleeping
for 6 hours. Furthermore, our investigation revealed a noteworthy effect
modification regarding the influence of total cholesterol levels on
systolic blood pressure when adjusted for specific confounding variables
within the studied cohort, especially in relation to age.

Our study contributes to the existing body of knowledge by
substantiating the associations uncovered in previous research
concerning sleep patterns and systolic blood pressure
[@su2022association]. This correlation underscores the significance of
sleep parameters as influential factors in understanding blood pressure
dynamics [@fuchs2020high]. In this way, the study explores factors that bear substantial weight in understanding disease risk within the aging US population. Findings from this study could inform public health policies and interventions towards healthy diet as well as sleep patterns, personalized medicine, and healthy aging promotion.

While our study used a representative sample of the US population, missing data on variables of interest in analyses presented a significant challenge. Our results are based on complete case analysis where all individuals with missing data on the variables of interest were excluded. While complete case analysis ensures that the analysis is based on comprehensive and complete information for each participant, the notable reduction in sample size has potential to result in low statistical power and bias when the assumption of missing completely at random does not hold [@baraldi2010introduction]. Future studies can address this challenge through such methods as multiple
imputation techniques. This approach involves generating multiple
plausible sets of missing values based on observed data patterns [@baraldi2010introduction].
